Predictive modeling with cross validation

### 2017-08-16 model fitting/cross_validation.Rmd
cross validation, updated hours

```{r}
first <- coxph(Surv(time_until_answer, answered) ~ category + as.factor(new_user) + contain_unanswered + title_questionmark + text_length + text_contain_punct + as.factor(num_freq_tags) + as.factor(n_tags) + contain_answered + newline_ratio + text_all_lower + update + greeting + capital_text + prior_effort + gratitude + ampm + n_images + weekday + device_length + title_beginwh, data = train1)
AIC(first, k = 2) #65294.56

@@

# backward/forward step-wise selection with stepAIC function from MASS package
model2 <- MASS::stepAIC(first, direction = "both", k = 2)
# results
results2 <- coxph(Surv(time_until_answer, answered) ~ category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh, data = train1)
AIC(results2, k = 2) # 65277.47
s <- summary(results2)
nrow(s$coefficients)

@@

# adding interaction terms
# adding title_questionmark*title_beginwh
model2.o <- coxph(Surv(time_until_answer, answered) ~ category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + title_questionmark*title_beginwh, data = train1)
AIC(model2.o, k = 2) # 65274.49
oshitar::compare_nested(model2.o, results2)

# adding new_user*update
model2.o1 <- coxph(Surv(time_until_answer, answered) ~ category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + new_user*update, data = train1)
AIC(model2.o1, k = 2) # 65272.55
oshitar::compare_nested(model2.o1, results2)

# adding new_user*ampm 
model2.o2 <- coxph(Surv(time_until_answer, answered) ~ category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + new_user*ampm, data = train1)
AIC(model2.o2, k = 2) # 65271.84
oshitar::compare_nested(model2.o2, results2)

@@

train1 %>%
  filter(new_category == "Electronics") %>%
  group_by(subcategory) %>%
  summarise(median = median(time_until_answer), n = n(), prop_a = sum(answered)/n) %>%
  arrange(median)

train1 %>%
  group_by(new_category) %>%
  summarise(median = median(time_until_answer), n = n(), prop_a = sum(answered)/n) %>%
  arrange(median)

@@

#new_cateogry*new_user
#new_cateogry*
model_new1 <- coxph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + new_category*new_user, data = train1)
AIC(model_new1, k = 2) # 64994.49
oshitar::compare_nested(model_new1, model_new)

@@

* Martingale residuals
    + plot for each continuous predictor 
    + assess adequacy of functional form of predictors, or suggest a potential form 
* Deviance residuals
    + check for outliers identify questions poorly predicted by the model 
    + one plot, examine questions with |residuals| > 2.5
* Schoenfeld residuals
    + one plot for each predictor 
    + if loess curve is flat, proportional hazard assumption is not violated, doesn't depend on time. Distinct pattern indicates violation of this assumption
    + **from the plots:** new_user, title_questionmark, num_freq_tags (1), ampm (Morning, Night), weekday (Thursday, Saturday), new_user*AMPM (morning, night) may violate PH assumption 
* Score residuals
    + check for influential observations. ID quesitons who's predictor values highly effect parameter estimates
    + plot for each predictor 
* **figure out how to identify certain points on ggplots**
* statified on new_category, AIC dropped drastically. But this means that I can't compare the different groups in new_category 
```{r}
#=============================================
#martingale residuals
train1$mart <- residuals(model_full, type = "martingale")

@@

@@

##### variable set up in test data

@@

```{r}
test1 <- x[splitPlan[[1]]$app, ]

#setting up variables in test1
#=============================================
test1$category <- as.character(test1$category)
test1$category[is.na(test1$category)] <- "Other"
#new_category
test1$new_category <- test1$category
apple_terms <- c("apple", "ipod", "ipad")
test1$apple <- str_detect(str_to_lower(test1$device), pattern = START %R% or1(apple_terms) %R% SPC)
test1$new_category[test1$apple == TRUE | test1$subcategory == "iPhone" | test1$category == "Mac"] <- "Apple Product"
test1$new_category[test1$new_category == "Phone"] <- "Android/Other Phone"
test1$new_category[test1$new_category == "Appliance" | test1$new_category == "Household"] <- "Home"
test1$new_category[test1$new_category == "Car and Truck" | test1$new_category == "Vehicle"] <- "Vehicle"
test1$new_category[test1$new_category == "Computer Hardware" | test1$new_category == "Media Player" | test1$new_category == "Apparel"] <- "Other"
# counts <- test1 %>%
#   group_by(new_category) %>%
#   summarise(n = n(), prop = n/nrow(test1))
# for (i in which(counts$prop <= 0.01)) {
#   test1$new_category[test1$new_category == counts$new_category[i]] <- "Other"
# }
#=============================================
#weekday
test1$datetime <- as.POSIXct(test1$post_date,origin="1970-01-01")
test1$weekday <- factor(weekdays(test1$datetime), levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
#=============================================
#ampm
test1$hour <- as.numeric(format(test1$datetime,"%H"))
test1$ampm <- "Night"
test1$ampm[test1$hour >= 5 & test1$hour < 12] <- "Morning"
test1$ampm[test1$hour >= 12 & test1$hour < 17] <- "Afternoon" #noon - 5pm
test1$ampm[test1$hour >= 17 & test1$hour < 20] <- "Evening" #5pm - 8pm
#=============================================
# text length
test1$text_length <- str_length(test1$text)
#=============================================
# device name length
test1$device_length <- str_length(test1$device)
#=============================================
# if the title ends with a question mark
library(rebus)
test1$title_questionmark <- str_detect(test1$title, pattern = QUESTION %R% END)
#=============================================
# if title begins with "wh"
test1$title_beginwh <- str_detect(str_to_lower(test1$title), pattern = "^wh")
#=============================================
test1$capital_text <- str_detect(as.character(test1$text), pattern = "^[[:upper:]]")
#=============================================
test1$update <- str_detect(test1$text, pattern = "===")
#=============================================
test1$newline_ratio <- str_count(test1$text, pattern = "\n")/str_length(test1$text)
#=============================================
# frequent tags
split_tags <- str_split(test1$tags, ", ", simplify = TRUE)
tag_vector <- as.vector(split_tags)
tag_vector <- tag_vector[which(tag_vector != "")]
unique_tags <- unique(tag_vector)

tag_freq <- data.frame(tag = unique_tags, percent = purrr::map_dbl(unique_tags, ~mean(rowSums(split_tags == .) > 0)))
tag_freq <- tag_freq %>%
              arrange(desc(percent))

#creating average frequency score variable
test1$tag1 <- split_tags[,1]
test1$tag2 <- split_tags[,2]
test1$tag3 <- split_tags[,3]
test1$tag4 <- split_tags[,4]

assign_score <- function(data, variable) {
  score <- rep(0, nrow(data))
  notempty <- which(data[[variable]] != "")
  for (i in notempty) {
    score[i] <- tag_freq$percent[which(tag_freq$tag == data[[variable]][i])]
  }
  return(score)
}
test1$score1 <- assign_score(test1, "tag1")
test1$score2 <- assign_score(test1, "tag2")
test1$score3 <- assign_score(test1, "tag3")
test1$score4 <- assign_score(test1, "tag4")
test1$avg_tag_score <- (test1$score1 + test1$score2 + test1$score3 + test1$score4)/as.numeric(test1$n_tags)
test1$avg_tag_score[is.nan(test1$avg_tag_score)] <- 0
#=============================================
# if question contains "frequent" tag
percentile80 <- quantile(test1$avg_tag_score, probs = 0.80)

test1$frequent_tag <- FALSE
test1$frequent_tag[test1$avg_tag_score >= percentile80] <- TRUE
#=============================================
#frequent terms in unanswered/answered questions
answered <- test1 %>%
  tbl_df() %>% 
  filter(answered == 1)
unanswered <- test1 %>%
  tbl_df() %>% 
  filter(answered == 0)

library(qdap)
library(tm)
terms_a <- oshitar::get_freq_terms(answered$title, stopwords = c("can", "will", "cant", "wont", "works", "get", "help", "need", "fix"))
terms_a$prop_in_answered <- terms_a$frequency/nrow(terms_a)
colnames(terms_a)[2] <- "frequency_a"

terms_u <- oshitar::get_freq_terms(unanswered$title, stopwords = c("can", "will", "cant", "wont", "works", "get", "help", "need", "fix"))
terms_u$prop_in_unanswered <- terms_u$frequency/nrow(terms_u)
colnames(terms_u)[2] <- "frequency_u"

combined <- dplyr::full_join(terms_a, terms_u, by = "word")
combined$ratio <- combined$prop_in_answered / combined$prop_in_unanswered

p_threshold <- 0.01
ratio_threshold <- 1

freq_terms_u <- combined %>%
                  filter(prop_in_unanswered > p_threshold) %>%
                  filter(ratio < ratio_threshold)
freq_terms_a <- combined %>%
                  filter(prop_in_answered > p_threshold) %>%
                  filter(ratio > ratio_threshold)

test1$contain_unanswered <- str_detect(as.character(test1$title), pattern = or1(freq_terms_u$word))
test1$contain_answered <- str_detect(as.character(test1$title), pattern = or1(freq_terms_a$word))

test1$new_user <- as.factor(test1$new_user)

#=============================================
# predicting on the test data

#356 NAs in the predictions
test_predict <- predict(model_full, newdata = test1, type = "risk", se.fit = TRUE)
test1$predictions <- test_predict[[1]]

@@

```{r}
# function to set up variables for test data 
variable_setup <- function(data) {
  data$category <- as.character(data$category)
  data$category[is.na(data$category)] <- "Other"

  data$new_category <- data$category
  apple_terms <- c("apple", "ipod", "ipad")
  data$apple <- str_detect(str_to_lower(data$device), pattern = START %R% or1(apple_terms) %R% SPC)
  data$new_category[data$apple == TRUE | data$subcategory == "iPhone" | data$category == "Mac"] <- "Apple Product"
  data$new_category[data$new_category == "Phone"] <- "Android/Other Phone"
  data$new_category[data$new_category == "Appliance" | data$new_category == "Household"] <- "Home"
  data$new_category[data$new_category == "Car and Truck" | data$new_category == "Vehicle"] <- "Vehicle"
  data$new_category[data$new_category == "Computer Hardware" | data$new_category == "Media Player" |
                     data$new_category == "Apparel"] <- "Other"
  #=============================================
  #weekday
  data$datetime <- as.POSIXct(data$post_date,origin="1970-01-01")
  data$weekday <- factor(weekdays(data$datetime), levels = c("Monday", "Tuesday", "Wednesday", 
                                                             "Thursday", "Friday", "Saturday", "Sunday"))
  #=============================================
  #ampm
  data$hour <- as.numeric(format(data$datetime,"%H"))
  data$ampm <- "Night"
  data$ampm[data$hour >= 5 & data$hour < 12] <- "Morning"
  data$ampm[data$hour >= 12 & data$hour < 17] <- "Afternoon" #noon - 5pm
  data$ampm[data$hour >= 17 & data$hour < 20] <- "Evening" #5pm - 8pm
  #=============================================
  # text length
  data$text_length <- str_length(data$text)
  #=============================================
  # device name length
  data$device_length <- str_length(data$device)
  #=============================================
  # if the title ends with a question mark
  library(rebus)
  data$title_questionmark <- str_detect(data$title, pattern = QUESTION %R% END)
  #=============================================
  # if title begins with "wh"
  data$title_beginwh <- str_detect(str_to_lower(data$title), pattern = "^wh")
  #=============================================
  data$capital_text <- str_detect(as.character(data$text), pattern = "^[[:upper:]]")
  #=============================================
  data$update <- str_detect(data$text, pattern = "===")
  #=============================================
  data$newline_ratio <- str_count(data$text, pattern = "\n")/str_length(data$text)
  #=============================================
  # frequent tags
  split_tags <- str_split(data$tags, ", ", simplify = TRUE)
  tag_vector <- as.vector(split_tags)
  tag_vector <- tag_vector[which(tag_vector != "")]
  unique_tags <- unique(tag_vector)

  tag_freq <- data.frame(tag = unique_tags, percent = purrr::map_dbl(unique_tags, ~mean(rowSums(split_tags == .) > 0)))
  tag_freq <- tag_freq %>%
              arrange(desc(percent))

  #creating average frequency score variable
  data$tag1 <- split_tags[,1]
  data$tag2 <- split_tags[,2]
  data$tag3 <- split_tags[,3]
  data$tag4 <- split_tags[,4]

  assign_score <- function(data, variable) {
    score <- rep(0, nrow(data))
    notempty <- which(data[[variable]] != "")
    for (i in notempty) {
      score[i] <- tag_freq$percent[which(tag_freq$tag == data[[variable]][i])]
    }
    return(score)
  }
  data$score1 <- assign_score(data, "tag1")
  data$score2 <- assign_score(data, "tag2")
  data$score3 <- assign_score(data, "tag3")
  data$score4 <- assign_score(data, "tag4")
  data$avg_tag_score <- (data$score1 + data$score2 + data$score3 + 
                            data$score4)/as.numeric(data$n_tags)
  data$avg_tag_score[is.nan(data$avg_tag_score)] <- 0
  #=============================================
  # if question contains "frequent" tag
  percentile80 <- quantile(data$avg_tag_score, probs = 0.80)

  data$frequent_tag <- FALSE
  data$frequent_tag[data$avg_tag_score >= percentile80] <- TRUE
  #=============================================
  #frequent terms in unanswered/answered questions
  answered <- data %>%
      tbl_df() %>% 
      filter(answered == 1)
  unanswered <- data %>%
    tbl_df() %>% 
    filter(answered == 0)

  library(qdap)
  library(tm)
  terms_a <- oshitar::get_freq_terms(answered$title, stopwords = c("can", "will", "cant", "wont", "works", "get", "help", "need", "fix"))
  terms_a$prop_in_answered <- terms_a$frequency/nrow(terms_a)
  colnames(terms_a)[2] <- "frequency_a"

  terms_u <- oshitar::get_freq_terms(unanswered$title, stopwords = c("can", "will", "cant", "wont", "works", "get", "help", "need", "fix"))
  terms_u$prop_in_unanswered <- terms_u$frequency/nrow(terms_u)
  colnames(terms_u)[2] <- "frequency_u"

  combined <- dplyr::full_join(terms_a, terms_u, by = "word")
  combined$ratio <- combined$prop_in_answered / combined$prop_in_unanswered

  p_threshold <- 0.01
  ratio_threshold <- 1

  freq_terms_u <- combined %>%
                    filter(prop_in_unanswered > p_threshold) %>%
                    filter(ratio < ratio_threshold)
  freq_terms_a <- combined %>%
                    filter(prop_in_answered > p_threshold) %>%
                    filter(ratio > ratio_threshold)

  data$contain_unanswered <- str_detect(as.character(data$title), pattern = or1(freq_terms_u$word))
  data$contain_answered <- str_detect(as.character(data$title), pattern = or1(freq_terms_a$word))

  data$new_user <- as.factor(data$new_user)

  return(data)
}

>>>>

```{r}
train1 <- x[splitPlan[[1]]$train, ]

first <- coxph(Surv(time_until_answer, answered) ~ category + as.factor(new_user) + contain_unanswered + title_questionmark + text_length + text_contain_punct + as.factor(num_freq_tags) + as.factor(n_tags) + contain_answered + newline_ratio + text_all_lower + update + greeting + capital_text + prior_effort + gratitude + ampm + n_images + weekday + device_length + title_beginwh, data = train1)
AIC(first, k = 2) #65294.56

@@

# backward/forward step-wise selection with stepAIC function from MASS package
model2 <- MASS::stepAIC(first, direction = "both", k = 2)
# results
results2 <- coxph(Surv(time_until_answer, answered) ~ category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh, data = train1)
AIC(results2, k = 2) # 65277.47

@@

# adding interaction terms
# adding title_questionmark*title_beginwh
model2 <- coxph(Surv(time_until_answer, answered) ~ category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + title_questionmark*title_beginwh, data = train1)
AIC(model2, k = 2) # 65274.49
compare_nested(model2, results2)

# adding new_user*update
model2_1 <- coxph(Surv(time_until_answer, answered) ~ category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + new_user*update, data = train1)
AIC(model2_1, k = 2) # 65272.55
compare_nested(model2_1, results2)

# adding new_user*ampm 
model2_2 <- coxph(Surv(time_until_answer, answered) ~ category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + new_user*ampm, data = train1)
AIC(model2_2, k = 2) # 65271.84
compare_nested(model2_2, results2)

@@

@@

#new_cateogry*new_user
#new_cateogry*
model_new1 <- coxph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + new_category*new_user, data = train1)
AIC(model_new1, k = 2) # 64994.49
compare_nested(model_new1, model_new)

@@

* Martingale residuals
    + plot for each continuous predictor 
    + assess adequacy of functional form of predictors, or suggest a potential form 
* Deviance residuals
    + check for outliers identify questions poorly predicted by the model 
    + one plot, examine questions with |residuals| > 2.5
* Schoenfeld residuals
    + one plot for each predictor 
    + if loess curve is flat, proportional hazard assumption is not violated, doesn't depend on time. Distinct pattern indicates violation of this assumption
    + **from the plots:** new_user, title_questionmark, num_freq_tags (1), ampm (Morning, Night), weekday (Thursday, Saturday), new_user*AMPM (morning, night) may violate PH assumption 
* Score residuals
    + check for influential observations. ID quesitons who's predictor values highly effect parameter estimates
    + plot for each predictor 
* **figure out how to identify certain points on ggplots**
* statified on new_category, AIC dropped drastically. But this means that I can't compare the different groups in new_category

```{r}
#=============================================
#martingale residuals
train1$mart <- residuals(model_full, type = "martingale")

@@

### Performance metric on training data set

```{r}
train_predict <- predict(object = final, type = "risk", se.fit = TRUE)
train1$predictions <- train_predict[[1]]
performance <- coxph(Surv(time_until_answer, answered) ~ predictions, data = train1)
summary(performance)
AIC(performance, k = 2)

#using cph from rms package to get Dxy and other performance metrics
library(rms)
options(scipen=999)
(cox <- cph(Surv(time_until_answer, answered) ~ predictions, data = train1, x = TRUE, y = TRUE, se.fit = TRUE, residuals = TRUE))
cox$stats
```

@@

@@

```{r}

#calculating root

test1 <- x[splitPlan[[1]]$app, ]
# setting up variables for model 
test1 <- oshitar::variable_setup(test1)

test_predict <- predict(object = final, newdata = test1, type = "risk", se.fit = TRUE)
test1$predictions <- test_predict[[1]]

@@

### working with the rms package for cross validation

```{r}
library(rms)
(cox <- cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + title_questionmark*title_beginwh + new_user*update + new_user*ampm + new_category*new_user, data = train1, x = TRUE, y = TRUE, se.fit = TRUE, residuals = TRUE))

(cv <- validate(cox, method = "crossvalidation", B = 10, dxy = TRUE, pr = TRUE))
```
